---
title: "Appendix Diet Summaries"
author: "Catalina Burch"
date: "2023-05-22"
classoption: landscape
output:
  pdf_document: default
---

```{r setup, include=FALSE}
#Load Libraries
library(here) #for finding working directory
library(readr) #for loading CSV
library(tidyverse)

#Load Data
focal <- read_csv(here("data/data.csv"))
allpred <- read_csv(here("data/all_pred_data.csv"))

```

## Appendix 1: Focal Predator Diet Compositions over Time

An additional component of this research project was to create data visualizations of diet summaries. These plots are intended to fill knowledge gaps in management documents, including the GOA Ecosystem Status Report and the ecosystem components of  commercial predators Stock Status Reports. Appendix one visualizes the change in the percent weight of prey contributions to diet over time for the four focal predator species at each size class. The size class is indicated at the top of each panel in black text. Red asterisks above bars indicates that there were fewer than 30 samples collected for that species/year/size class. 

\newpage
```{r A1Plot1, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
APX1 <- focal %>% 
  group_by(Pred_common, Len_bin_sample, Year) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Pred_common, Len_bin_sample, stock_groupings, Year) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin_sample, stock_groupings, Year, TotalWt,PreyWt, PW, n)

#ordering the length bins
APX1$Len_bin_sample <- factor(APX1$Len_bin_sample, 
                            levels = c("<25", "<29", "<39",
                                       "25-39", "29-44", "29-49", "39-69", "40-54", "45-59",
                                       ">39", ">49", ">54", ">59", ">69"))
unique(APX1$Len_bin_sample)

#Colors for plot
colorlist<-c('#8A0000', "#C80000", '#e34a33', '#fc8d59', '#fdcc8a', '#fef0d9', 'white', 
                      '#f1eef6', '#d0d1e6', '#a6bddb', '#74a9cf', '#3690c0', '#0570b0', '#034e7b')
                      
#Ordering the prey items
APX1$stock_groupings <- factor(APX1$stock_groupings, 
                                 levels = c("Arthropoda", "commercial crab", "copepod", "benthic invertebrate", "euphausiids",
                                            "zooplankton", "other","cod", "flatfish",
                                            "walleye pollock","salmon","rockfish", "forage fish", "other fish"))

#Appendix Plot 1
APXPlot1 <- APX1 %>% 
  filter(Pred_common == "Arrowtooth flounder") %>% 
  ggplot(aes(x = Year, y = PW, fill = stock_groupings)) + 
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Year, y = 103, label = ifelse(n<=30, "*", "")), color = '#b30000') +
  facet_wrap(~Len_bin_sample) +
  scale_fill_manual(values = colorlist, name = "Prey Groupings") +
  labs(x = "Year", y = "Percent weight ", title = "Arrowtooth Flounder Diet") +
  scale_y_continuous(expand = c(0,4)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30))

APXPlot1
```

Plot description
\newpage
```{r A1Plot2, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Appendix Plot 2
APXPlot2 <- APX1 %>% 
  filter(Pred_common == "Pacific halibut") %>% 
  ggplot(aes(x = Year, y = PW, fill = stock_groupings)) + 
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Year, y = 103, label = ifelse(n<=30, "*", "")), color = '#b30000') +
  facet_wrap(~Len_bin_sample) +
  scale_fill_manual(values = colorlist, name = "Prey Groupings") +
  labs(x = "Year", y = "Percent weight ", title = "Pacific halibut Diet") +
  scale_y_continuous(expand = c(0,4)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30))

APXPlot2
```

Plot description

```{r A1Plot3, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
APXPlot3 <- APX1 %>% 
  filter(Pred_common == "Walleye pollock") %>% 
  ggplot(aes(x = Year, y = PW, fill = stock_groupings)) + 
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Year, y = 103, label = ifelse(n<=30, "*", "")), color = '#b30000') +
  facet_wrap(~Len_bin_sample) +
  scale_fill_manual(values = colorlist, name = "Prey Groupings") +
  labs(x = "Year", y = "Percent weight ", title = "Walleye pollock Diet") +
  scale_y_continuous(expand = c(0,4)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30))

APXPlot3
```

Plot description

```{r A1Plot4, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Appendix Plot 4
APXPlot4 <- APX1 %>% 
  filter(Pred_common == "Pacific cod") %>% 
  ggplot(aes(x = Year, y = PW, fill = stock_groupings)) + 
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Year, y = 103, label = ifelse(n<=30, "*", "")), color = '#b30000') +
  facet_wrap(~Len_bin_sample) +
  scale_fill_manual(values = colorlist, name = "Prey Groupings") +
  labs(x = "Year", y = "Percent weight ", title = "Pacific cod Diet") +
  scale_y_continuous(expand = c(0,4)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30))

APXPlot4
```

Plot description

\newpage

## Appendix 2: Focal Prey Occurrence and PSIRI Over Time

Write out a description of this section

\newpage
```{r A2Plot1A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
APX2 <- focal %>% 
  group_by(Pred_common, Len_bin_sample, Year) %>% 
  mutate(totn = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Pred_common, Len_bin_sample, Prey_Name, Year) %>% 
  mutate(preyn = length(unique(uniqueID)),
         PO = (preyn/totn)*100) %>% 
  distinct(Pred_common, Len_bin_sample, Prey_Name, totn, Year, PO)

APX2$Len_bin_sample <- factor(APX2$Len_bin_sample, 
                              levels = c("<25", "<29", "<39",
                                         "25-39", "29-44", "29-49", "39-69", 
                                         "40-54", "45-59",
                                         ">49", ">54", ">59", ">69"))

levels(APX2$Len_bin_sample) = c("1", "1", "1",
                                  "2", "2", "2", "2", 
                                  "3", "3",
                                  "3", "4", "4", "3")



colorlist2<-c('#a6bddb', '#3690c0', '#0570b0', '#034e7b')

#Plot 1: Euphausiids
APX2Plot1 <- APX2 %>% 
  filter(Prey_Name == "Euphausiacea") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Euphausiid Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot1
```

Plot description


```{r A2Plot1B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
APX3 <- focal %>% 
  drop_na(PREY_CNT, PREY_TWT) %>% 
  group_by(Pred_common, Len_bin_sample, Year) %>% 
  mutate(TotalWt = sum(PREY_TWT), totn = length(unique(uniqueID)), count = sum(PREY_CNT)) %>% 
  ungroup() %>% 
  group_by(Pred_common, Len_bin_sample, Prey_Name, Year) %>% 
  mutate(PreyWt = sum(PREY_TWT), preyn = length(unique(uniqueID)), preycount = sum(PREY_CNT),
         PW = (PreyWt/TotalWt)*100, PO = (preyn/totn)*100, PN = (preycount/count)*100,
         PSIRI = (PO*(PN+PW))/2) %>% 
  distinct(Pred_common, Len_bin_sample, stock_groupings, Year, totn, PSIRI)

APX3$Len_bin_sample <- factor(APX3$Len_bin_sample, 
                                levels = c("<25", "<29", "<39",
                                           "25-39", "29-44", "29-49", "39-69", 
                                           "40-54", "45-59",
                                           ">49", ">54", ">59", ">69"))

levels(APX3$Len_bin_sample) = c("1", "1", "1",
                                  "2", "2", "2", "2", 
                                  "3", "3",
                                  "3", "4", "4", "3")

#Plot 1: Euphausiids
APX3Plot1 <- APX3 %>% 
  filter(Prey_Name == "Euphausiacea") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Euphausiid PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot1
```

Plot Description


```{r A2Plot2A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 2: Pollock
APX2Plot2 <- APX2 %>% 
  filter(Prey_Name == "Walleye pollock") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Walleye Pollock Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot2
```

Plot Description

```{r A2Plot2B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 2: Pollock
APX3Plot2 <- APX3 %>% 
  filter(Prey_Name == "Walleye pollock") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Walleye Pollock PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot2
```

Plot Description

```{r A2Plot3A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 3: Osmerids
APX2Plot3 <- APX2 %>% 
  filter(Prey_Name == "Osmerid") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Osmeridae Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot3
```

Plot Description

```{r A2Plot3B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 3: Osmerids
APX3Plot3 <- APX3 %>% 
  filter(Prey_Name == "Osmerid") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Osmeridae PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot3
```

Plot Description

```{r A2Plot4A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 4: Ammodytidae
APX2Plot4 <- APX2 %>% 
  filter(Prey_Name == "Ammodytidae") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Ammodytidae Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot4
```

Plot Description

```{r A2Plot4B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 4: Ammodytidae
APX3Plot4 <- APX3 %>% 
  filter(Prey_Name == "Ammodytidae") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Ammodytidae PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot4
```

Plot Description

```{r A2Plot5A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 5: Clupeidae
APX2Plot5 <- APX2 %>% 
  filter(Prey_Name == "Clupeoidei") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Clupeidae Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot5
```

Plot Description

```{r A2Plot5B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 5: Clupeidae
APX3Plot5 <- APX3 %>% 
  filter(Prey_Name == "Clupeoidei") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Clupeidae PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot5
```

Plot Description

```{r A2Plot6A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 6: Paguridae
APX2Plot6 <- APX2 %>% 
  filter(Prey_Name == "Paguridae") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Paguridae Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot6
```

Plot Description

```{r A2Plot6B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 6: Paguridae
APX3Plot6 <- APX3 %>% 
  filter(Prey_Name == "Paguridae") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Paguridae PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot6
```

Plot Description

```{r A2Plot7A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 7: Pandalidae
APX2Plot7 <- APX2 %>% 
  filter(Prey_Name == "Pandalidae (shrimp)") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Pandalidae Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot7
```

Plot Description

```{r A2Plot7B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 7: Pandalidae
APX3Plot7 <- APX3 %>% 
  filter(Prey_Name == "Pandalidae (shrimp)") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Pandalidae PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot7
```

Plot Description

```{r A2Plot8A, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 8: Tanner Crab
APX2Plot8 <- APX2 %>% 
  filter(Prey_Name == "Tanner Crab") %>% 
  ggplot(aes(x = Year, y = PO, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Tanner Crab Occurrence in Diets", y = "Percent Occurrence", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX2Plot8
```

Plot Description

```{r A2Plot8B, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Plot 8: Tanner Crab
APX3Plot8 <- APX3 %>% 
  filter(Prey_Name == "Tanner Crab") %>% 
  ggplot(aes(x = Year, y = PSIRI, color = Len_bin_sample)) +
  geom_line() +
  geom_point(aes(size = totn), alpha = 0.5) +
  facet_wrap(~Pred_common)+
  scale_color_manual(values = colorlist2, name = "Predator length bin") +
  theme_minimal() +
  labs(title = "Tanner Crab PSIRI in Diets", y = "PSIRI", size = "Sample size") +
  theme(axis.text.x = element_text(angle = 30))

APX3Plot8
```

Plot Description

\newpage

## Appendix 3: Non-focal Predators General Diet Summaries

Write up a description of this section

\newpage

```{r A3Plot1, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#SABLEFISH
#calculate percent weight and sample size based on sablefish specific prey groups
sable <- allpred %>% 
  filter(Pred_common == "Sablefish") %>% 
  mutate(Len_bin = cut(PRED_LEN, breaks = c(0, 40, 50, 60, 70, 280))) %>% 
  group_by(Len_bin) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Len_bin, sable_group) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin, sable_group, PW, n)
levels(sable$Len_bin) <- c("<40", "40-50", "51-60", "61-70", ">70")

#colors for plot
colorlistS <-c('#8A0000', '#fc8d59', '#fdcc8a', '#fef0d9', 
                        '#c51b8a', '#fa9fb5', '#fde0dd',
                      'white', 
                      '#f1eef6','#a6bddb', '#0570b0', '#034e7b', 'black')

#Ordering the prey items
sable$sable_group <- factor(sable$sable_group, 
                               levels = c("Arthropoda", "benthic invertebrate", "euphausiids", "zooplankton", 
                                          "squid", "jellyfish","octopus",
                                          "other",
                                          "cod", "walleye pollock", "forage fish", "other fish", "offal"))

#Plot diet figure
APX4Plot1 <- sable %>% 
  ggplot( aes(x = Len_bin, y = PW, fill = sable_group)) +
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Len_bin, y = 103, label = n), color = '#b30000') +
  scale_fill_manual(values = colorlistS, name = "Prey Groupings") +
  labs(x = "Predator length (cm)", y = "Percent weight", title = "Sablefish Diet") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = c(0,2)) +
  theme_minimal() 

APX4Plot1
```
Plot description

```{r A3Plot2, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Atka mackerel
#calculate percent weight and sample size based on sablefish specific prey groups
atka <- allpred %>% 
  filter(Pred_common == "Atka mackerel") %>% 
  mutate(Len_bin = cut(PRED_LEN, breaks = c(0, 40, 280))) %>% 
  group_by(Len_bin) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Len_bin, atka_group) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin, atka_group, PW, n)
levels(atka$Len_bin) <- c("<40", ">40")

#colors for plot
colorlistA <-c('#e34a33', '#fdcc8a', '#fef0d9',
                        'white')
                        
#Ordering the prey items
atka$atka_group <- factor(atka$atka_group, 
                            levels = c("copepod", "euphausiids", "zooplankton", 
                                       "other"))

#Plot diet figure
APX4Plot2 <- atka %>% 
  ggplot( aes(x = Len_bin, y = PW, fill = atka_group)) +
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Len_bin, y = 103, label = n), color = '#b30000') +
  scale_fill_manual(values = colorlistA, name = "Prey Groupings") +
  labs(x = "Predator length (cm)", y = "Percent weight", title = "Atka Mackerel Diet") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = c(0,2)) +
  theme_minimal() 

APX4Plot2
```
Plot description

```{r A3Plot3, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Pacific Ocean Perch
#calculate percent weight and sample size based on sablefish specific prey groups
POP <- allpred %>% 
  filter(Pred_common == "Pacific ocean perch") %>% 
  mutate(Len_bin = cut(PRED_LEN, breaks = c(0, 20, 30, 40, 280))) %>% 
  group_by(Len_bin) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Len_bin, POP_grouping) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin, POP_grouping, PW, n)
levels(POP$Len_bin) <- c("<20","20-30", "31-40", ">40")

#Colors for plot
colorlistP<-c('#8A0000', '#e34a33', '#fc8d59', '#fdcc8a', '#fef0d9', 'white', 
                       '#a6bddb',  '#034e7b')
                      
#Ordering the prey items
POP$POP_grouping <- factor(POP$POP_grouping, 
                               levels = c("Arthropoda", "copepod", "benthic invertebrate", "euphausiids","zooplankton",
                                          "other",
                                          "walleye pollock","salmon","rockfish", "forage fish", "other fish"))

#Plot diet figure
APX4Plot3 <- POP %>% 
  ggplot( aes(x = Len_bin, y = PW, fill = POP_grouping)) +
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Len_bin, y = 103, label = n), color = '#b30000') +
  scale_fill_manual(values = colorlistP, name = "Prey Groupings") +
  labs(x = "Predator length (cm)", y = "Percent weight", title = "Pacific Ocean Perch Diet") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = c(0,2)) +
  theme_minimal() 

APX4Plot3
```

Plot Description

```{r A3Plot4, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Flathead Sole
#calculate percent weight and sample size based on sablefish specific prey groups
Flathead <- allpred %>% 
  filter(Pred_common == "Flathead sole") %>% 
  mutate(Len_bin = cut(PRED_LEN, breaks = c(0, 20, 30, 40, 280))) %>% 
  group_by(Len_bin) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Len_bin, flathead_groupings) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin, flathead_groupings, PW, n)
levels(Flathead$Len_bin) <- c("<20","20-30", "31-40", ">40")

#Colors for plot
colorlistF<-c('#8A0000', "#C80000", '#fc8d59', '#fdcc8a', '#fef0d9', 
                       "#7a0177", "#c51b8a", "#f768a1", "#fa9fb5", "#fcc5c0", 
                       'white', 
                       '#034e7b')
                       
#Ordering the prey items
Flathead$flathead_groupings <- factor(Flathead$flathead_groupings, 
                           levels = c("Arthropoda","commercial crab", "benthic invertebrate", "euphausiids","zooplankton",
                                      "Bivalvia", "brittle star", "Crangonidae", "Paguridae", "Pandalidae",
                                      "other",
                                      "other fish"))

#Plot diet figure
APX4Plot4 <- Flathead %>% 
  ggplot( aes(x = Len_bin, y = PW, fill = flathead_groupings)) +
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Len_bin, y = 103, label = n), color = '#b30000') +
  scale_fill_manual(values = colorlistF, name = "Prey Groupings") +
  labs(x = "Predator length (cm)", y = "Percent weight", title = "Flathead Sole Diet") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = c(0,2)) +
  theme_minimal() 

APX4Plot4
```

Plot description

```{r A3Plot5, echo=FALSE, results=FALSE, fig.width= 9, fig.height=5, dpi = 300}
#Southern Rock Sole
#calculate percent weight and sample size based on specific prey groups
SRS <- allpred %>% 
  filter(Pred_common == "Southern rock sole") %>% 
  mutate(Len_bin = cut(PRED_LEN, breaks = c(0, 20, 30, 40, 280))) %>% 
  group_by(Len_bin) %>% 
  mutate(TotalWt = sum(PREY_TWT), n = length(unique(uniqueID))) %>% 
  ungroup() %>% 
  group_by(Len_bin, srs_groupings) %>% 
  mutate(PreyWt = sum(PREY_TWT), PW = (PreyWt/TotalWt)*100) %>% 
  distinct(Pred_common, Len_bin, srs_groupings, PW, n)
levels(Flathead$Len_bin) <- c("<20","20-30", "31-40", ">40")

#Colors for plot
colorlistS<-c('#8A0000', '#fc8d59', '#fef0d9',
                      "#7a0177", "#c51b8a", "#fcc5c0", 
                      'white', 
                      '#d0d1e6', '#0570b0', '#034e7b')
                      
#Ordering the prey items
SRS$srs_groupings <- factor(SRS$srs_groupings, 
                                 levels = c("Arthropoda", "benthic invertebrate", "zooplankton",
                                            "Bivalvia", "brittle star", "Polychaeta",
                                            "other",
                                            "flatfish", "forage fish", "other fish"))

#Plot diet figure
APX4Plot5 <- SRS %>% 
  ggplot( aes(x = Len_bin, y = PW, fill = srs_groupings)) +
  geom_bar(stat = "identity", show.legend = T) +  
  geom_text(aes(x = Len_bin, y = 103, label = n), color = '#b30000') +
  scale_fill_manual(values = colorlistS, name = "Prey Groupings") +
  labs(x = "Predator length (cm)", y = "Percent weight", title = "Southern Rock Sole Diet") +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_y_continuous(expand = c(0,2)) +
  theme_minimal() 

APX4Plot5

```

Plot description