---
title: "GAM Models"
author: "Catalina Burch"
date: "2023-05-24"
output: pdf_document
---

This document runs all of the GAMs included in my thesis.

```{r setup, include=FALSE}
library(here) #for finding working directory
library(readr) #for loading CSV
library(writexl) #for exporting 
library(tidyverse)
library(mgcv) #for running gams
library(MuMIn) #for the dredge summary table
library(patchwork) #for combining plots
library(gridExtra) #for laying out plots
library(visreg) #for visualizing partial effects
library(mapdata) #for partial effects map
library(gamm4)
library(mgcViz)
options(na.action = "na.fail") 

#load data
Mdata <- read_csv(here("data/data.csv")) #this data includes empty stomachs

#change year to factor
Mdata$Year <- factor(Mdata$Year)
Mdata$MHW <- factor(Mdata$MHW)
Mdata$STATIONID <- factor(Mdata$STATIONID)

#Change to binary wide format data
Mdata <- Mdata %>% 
  mutate(pres_absent = 1) %>% #create binary presence for each prey item
    distinct(Year, uniqueID, Haul_Join, RLAT, RLONG, GEAR_DEPTH, MHW, STATIONID,
             GEAR_TEMP, PRED_LEN, Pred_common, Prey_Name, pres_absent) %>% #remove redundancies, i.e. the same prey species listed twice for the same stomach with different life history stages
    pivot_wider(names_from = Prey_Name, values_from = pres_absent, values_fill = list(pres_absent = 0)) %>% #create wide dataframe with a column for each prey type
    rename(Walleyepollock = 'Walleye pollock', TannerCrab = 'Tanner Crab',
           Crangonidae = 'Crangonidae (shrimp)', Pandalidae = `Pandalidae (shrimp)`) %>%  #rename (this was an issue for running the prey models below because the model doesn't like a space in the column name)
  group_by(uniqueID) %>% 
  mutate(forage = sum(Osmerid, Clupeoidei),
         forage = ifelse(forage > 0, 1, 0))

Mdata$MHW <- factor(Mdata$MHW, levels = c("Pre", "MHW", "Post"))

#create separate dataframes for each predator
WP <- Mdata %>% 
  filter(Pred_common == "Walleye pollock") 

PH <- Mdata %>% 
  filter(Pred_common == "Pacific halibut") 

PC <- Mdata %>% 
  filter(Pred_common == "Pacific cod") 

AF <- Mdata %>% 
  filter(Pred_common == "Arrowtooth flounder") 

# Set coordinate boundaries for plotting:
lonmin = -172
lonmax = -130
latmin = 52
latmax = 62

```

## Model 1: Euphausiacea Prey
```{r EuphWP, echo=FALSE, results=FALSE}
#Walleye Pollock
#MHW Interaction Model

start <- Sys.time()

Model1 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = WP,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Gamm4 package
# start <- Sys.time()
# Model1_gamm4 <- gamm4(Euphausiacea ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) + 
#                 s(PRED_LEN, k = 4, by = MHW),
#               random = ~(1|Year/STATIONID),
#     data = WP,
#     family = binomial())
# 
# end <- Sys.time()
# mod1.time <- end - start
# 
# gam.check(Model1$gam)
# p <- getViz(Model2$gam)
# 
# plot(p)
# 
# summary(Model1$gam)
# gam.vcomp(Model2$gam, conf.lev = .95)
# 
# library(gratia)
# variance_comp(Model1$gam)



#Base Model
start <- Sys.time()
Model2 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = WP,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat <- matrix(nrow = 16, ncol = 5)

AIC_Mat[1,1] <- "Prey"
AIC_Mat[1,2]<- "Pred"
AIC_Mat[1,3] <- "MHW Model AIC"
AIC_Mat[1,4] <- "Base Model AIC"
AIC_Mat[1,5] <- "Delta AIC"
AIC_Mat[2,1] <- "Euph"
AIC_Mat[2,2] <- "WP"
AIC_Mat[2,3] <- AIC(Model1)
AIC_Mat[2,4] <- AIC(Model2)
AIC_Mat[2,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Gamm4
# start <- Sys.time()
# Model2 <- gamm4(Euphausiacea ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4),
#               random = ~(1|Year/STATIONID),
#     data = WP,
#     family = binomial())
# 
# end <- Sys.time()
# mod2.time <- end - start
# 
# #Blog post https://fromthebottomoftheheap.net/2021/02/02/random-effects-in-gams/
# 
# AIC(Model1$mer) - AIC(Model2$mer)


#Plotting partial effects

#Model1
Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response", overlay = T,
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response", overlay = T,
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response", 
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response", 
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

#Model2
Plot3 <- visreg(Model2, "GEAR_TEMP", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot4 <- visreg(Model2, "PRED_LEN", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Walleye Pollock") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Walleye Pollock") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Euphausiacea Occurrence")

grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Euphausiacea Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)

# plot.new()
# map('worldHires', fill=T, xlim=c(-172, -130), ylim=c(52, 62), col="lightgrey") #I need to use this map twice to fix the aspect ratio.
# 
# plot(Model1)
# vis.gam(Model1, view = c("RLONG", "RLAT"), plot.type = "contour", type="response", 
#         contour.col="black", color="heat", xlab="Longitude", ylab="Latitude", 
#         main="", too.far=0.025, n.grid=250, 
#         xlim=c(lonmin, lonmax), ylim=c(latmin, latmax), add = T) 
# map('worldHires', fill=T, xlim=c(-170, -130), ylim=c(52, 62), col="lightgrey", add = T)

```

## Model 1: Euphausiacea Prey
```{r EuphPC, echo=FALSE, results=FALSE}
#Pacific Cod
#MHW Interaction Model

start <- Sys.time()

Model1 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[3,1] <- "Euph"
AIC_Mat[3,2] <- "PC"
AIC_Mat[3,3] <- AIC(Model1)
AIC_Mat[3,4] <- AIC(Model2)
AIC_Mat[3,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Cod") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Cod") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Euphausiacea Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Euphausiacea Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)

```


## Model 1: Euphausiacea Prey
```{r EuphAF, echo=FALSE, results=FALSE}
#Arrowtooth Flounder
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Euphausiacea ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[4,1] <- "Euph"
AIC_Mat[4,2] <- "AF"
AIC_Mat[4,3] <- AIC(Model1)
AIC_Mat[4,4] <- AIC(Model2)
AIC_Mat[4,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Euphausiacea Occurrence")

grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Euphausiacea Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```












## Model 2: Walleye Pollock Prey
```{r PollPH, echo=FALSE, results=FALSE}
#Pacific Halibut
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Mat[5,1] <- "WP"
AIC_Mat[5,2] <- "PH"
AIC_Mat[5,3] <- AIC(Model1)
AIC_Mat[5,4] <- AIC(Model2)
AIC_Mat[5,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Halibut") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Walleye Pollock Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Walleye Pollock Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```



## Model 2: Walleye Pollock Prey
```{r PollPC, echo=FALSE, results=FALSE}
#Pacific Cod
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[6,1] <- "WP"
AIC_Mat[6,2] <- "PC"
AIC_Mat[6,3] <- AIC(Model1)
AIC_Mat[6,4] <- AIC(Model2)
AIC_Mat[6,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "", 
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Cod") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Cod")

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Walleye Pollock Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Walleye Pollock Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```



## Model 2: Walleye Pollock Prey
```{r PollAF, echo=TRUE, results=FALSE}
#Arrowtooth Flounder
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Walleyepollock ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[7,1] <- "WP"
AIC_Mat[7,2] <- "AF"
AIC_Mat[7,3] <- AIC(Model1)
AIC_Mat[7,4] <- AIC(Model2)
AIC_Mat[7,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Arrowtooth Flouder") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Arrowtooth Flouder") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Walleye Pollock Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Walleye Pollock Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```









## Model 3: Pandalidae Prey
```{r PanWP, echo=FALSE, results=FALSE}
#Walleye Pollock
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = WP,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = WP,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[8,1] <- "Pandalid"
AIC_Mat[8,2] <- "WP"
AIC_Mat[8,3] <- AIC(Model1)
AIC_Mat[8,4] <- AIC(Model2)
AIC_Mat[8,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic()

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Walleye Pollock") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Walleye Pollock") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Pandalidae Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Pandalidae Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```



## Model 3: Pandalidae Prey
```{r PanPC, echo=FALSE, results=FALSE}
#Pacific Cod
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[9,1] <- "Pandalid"
AIC_Mat[9,2] <- "PC"
AIC_Mat[9,3] <- AIC(Model1)
AIC_Mat[9,4] <- AIC(Model2)
AIC_Mat[9,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Walleye Pollock") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Walleye Pollock") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Pandalidae Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Pandalidae Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)

```


## Model 3: Pandalidae Prey
```{r PanAF, echo=FALSE, results=FALSE}
#Arrowtooth Flounder
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Pandalidae ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[10,1] <- "Pandalid"
AIC_Mat[10,2] <- "AF"
AIC_Mat[10,3] <- AIC(Model1)
AIC_Mat[10,4] <- AIC(Model2)
AIC_Mat[10,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic()

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Pandalidae Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Pandalidae Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)

```










## Model 5: Osmerid Prey
```{r OsmPH, echo=FALSE, results=FALSE}
#Pacific Halibut
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Osmerid ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Osmerid ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[11,1] <- "Osmerid"
AIC_Mat[11,2] <- "PH"
AIC_Mat[11,3] <- AIC(Model1)
AIC_Mat[11,4] <- AIC(Model2)
AIC_Mat[11,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Halibut") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Osmerid Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Osmerid Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```



## Model 5: Osmerid Prey
```{r OsmAF, echo=FALSE, results=FALSE}
#Arrowtooth Flounder
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Osmerid ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Osmerid ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = AF,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[12,1] <- "Osmerid"
AIC_Mat[12,2] <- "AF"
AIC_Mat[12,3] <- AIC(Model1)
AIC_Mat[12,4] <- AIC(Model2)
AIC_Mat[12,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Arrowtooth Flounder") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Osmerid Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```







## Model 6: Tanner Crab Prey Occurrence
```{r TanPH, echo=FALSE, results=FALSE}
#Pacific Halibut
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(TannerCrab ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(TannerCrab ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[13,1] <- "TannerCrab"
AIC_Mat[13,2] <- "PH"
AIC_Mat[13,3] <- AIC(Model1)
AIC_Mat[13,4] <- AIC(Model2)
AIC_Mat[13,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Halibut") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Tanner Crab Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Tanner Crab Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```


## Model 6: Tanner Crab Prey Occurrence
```{r TanPC, echo=FALSE, results=FALSE}
#Pacific Cod
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(TannerCrab ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(TannerCrab ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[14,1] <- "TannerCrab"
AIC_Mat[14,2] <- "PC"
AIC_Mat[14,3] <- AIC(Model1)
AIC_Mat[14,4] <- AIC(Model2)
AIC_Mat[14,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Halibut") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Tanner Crab Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Tanner Crab Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```




## Model 7: Paguridae Prey Occurrence
```{r PagPH, echo=FALSE, results=FALSE}
#Pacific Halibut
#MHW Interaction Model
start <- Sys.time()

Model1 <- gam(Paguridae ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Paguridae ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PH,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[15,1] <- "Pagurid"
AIC_Mat[15,2] <- "PH"
AIC_Mat[15,3] <- AIC(Model1)
AIC_Mat[15,4] <- AIC(Model2)
AIC_Mat[15,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic()

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Halibut") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Paguridae Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Paguridae Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```


## Model 7: Paguridae Prey Occurrence
```{r PagPC, echo=FALSE, results=FALSE}
#Pacific Cod
start <- Sys.time()

Model1 <- gam(Paguridae ~  s(GEAR_TEMP, k = 4, by = MHW) + s(RLONG, RLAT, by = MHW) +
                s(PRED_LEN, k = 4, by = MHW) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod1.time <- end - start

#Base Model
start <- Sys.time()
Model2 <- gam(Paguridae ~  s(GEAR_TEMP, k = 4) + s(RLONG, RLAT) + s(PRED_LEN, k = 4) +
                s(Year, bs = "re"),
    data = PC,
    family = binomial())

end <- Sys.time()
mod2.time <- end - start

AIC_Diff <- AIC(Model1) - AIC(Model2)

AIC_Mat[16,1] <- "Pagurid"
AIC_Mat[16,2] <- "PC"
AIC_Mat[16,3] <- AIC(Model1)
AIC_Mat[16,4] <- AIC(Model2)
AIC_Mat[16,5] <- (AIC(Model1) - AIC(Model2))

print(AIC_Mat)

#Plotting partial effects

#Model1
Plot1 <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "") +
  theme_classic() 

Plot2 <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "") +
  theme_classic() 

Plot1_O <- visreg(Model1, "GEAR_TEMP", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Gear temperature", ylab = "",
       overlay = T) +
  theme_classic() 

Plot2_O <- visreg(Model1, "PRED_LEN", "MHW", type = "conditional", scale = "response",
       gg = TRUE, line=list(col="black"), xlab = "Predator Length", ylab = "",
       overlay = T) +
  theme_classic() 

MainP <- (Plot1 / Plot2) + 
  plot_annotation(title = "Predator: Pacific Cod") 
MainPO <- (Plot1_O / Plot2_O) + 
  plot_annotation(title = "Predator: Pacific Cod") 

grid.arrange(patchworkGrob(MainP), left = "Partial Effect on Paguridae Occurrence")
grid.arrange(patchworkGrob(MainPO), left = "Partial Effect on Paguridae Occurrence")

p <- getViz(Model1)
 
plot(p, select = c(4,5,6), pages = 1)
```
