---
title: "Walleye Pollock Summary Tables"
output: html_document
author: Catalina Burch
date (last edited): Aug 1, 2022
---

```{r setup, include=FALSE}
#load libraries
library(readr) 
library(tidyverse)
library(here)
library(taxize)

#load data
raw_stomach_contents <- read_csv(here("data/GOA_Raw_StomachContents.csv"))

```

##Calculating Percent Frequency of Occurance and Percent Weight

The %W (Percent Weight) for a given prey type i is calculated by summing the weight of prey type i and dividing by the total weight of all prey, then multiplying by 100. 

The %N (Percent Number) for a given prey type i is calculated by summing the number of individuals of prey type i and dividing by the total numerical abundance of all prey, then multiplying by 100. 

The %F (Percent Frequency of Occurrence) for a given prey type i is calculated as the number of lingcod that had at least one observation of prey i in their stomachs divided by the total number of lingcod sampled, multiplied by 100.

```{r Pollock}
#YEAR: 1999
#calculating percent weight and percent number
WN_poll_1999 <- raw_stomach_contents %>% 
  filter(Year == 1999, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_1999 <- raw_stomach_contents %>% 
  filter(Year == 1999, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 1999)

#I know that the total number of pollock stomachs sampled is 540 because of the summary table

#merge dataframes
poll_1999 <- full_join(WN_poll_1999, F_poll_1999, by = "Prey_Name")

#checking that the % sum to 100
sum(poll_1999$`%W`)
sum(poll_1999$`%N`)
sum(poll_1999$`%F`)
sum(poll_1999$F)

#%F does not sum to 100, but I think that is because for F I looked at the number of Pollock stomachs that contained at least one observation of each prey type, but there are likely stomachs that have multiple prey types within each stomach which would result in overcounting the sum(F)[764] compared to the total number of pollock stomachs sampled (540)

#I've also noticed that the numbers here do not match Table 3 from Yang et al. I'm wondering if it has to do with prey groupings, maybe the groupings have changed since that paper was published? Maybe it has something to do with empty stomachs, should I not include empty stomachs?

```

#This chunk is me trying to figure out how to write a loop.
```{r loop}
#I wrote a function to calculate %W  %N and %F
PW_PN_PF <- function(.data) {
  .data %>% 
  group_by(Prey_Name) %>% 
    summarise(TotalWt = sum(Prey_twt), N = n(), F = sum(unique(Pred_specn)*0+1)) %>% 
    mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100, "%F" = (F/540)*100)
}


#vector of variable to run through the loop
years <- c(1999, 2001, 2003, 2005, 2007, 2009, 2011, 2013, 2015, 2017, 2019)

#object to contain results
pollock_summary <- tibble()

#the for loop
for(i in years){
  pollock <- raw_stomach_contents %>% 
  filter(Pred_nodc == 8791030701, Year == i) %>% 
  PW_PN_PF() %>% 
    mutate(Year = i)
  
  pollock_summary <- bind_rows(pollock_summary, pollock)
}

```


```{r, figure}
ggplot(pollock_summary, aes(y = '%W', x = years)) +
  geom_bar(position = "fill", stat = "identity")

```


#THIS FOLLOWING CHUNK IS LIKELY OBSOLETE
#Next I will calculate this summary information for the other years of data
#This chunk of code may become obsolete after I figure out how to make an efficient loop. 
```{r }
#Year: 2001
WN_poll_2001 <- raw_stomach_contents %>% 
  filter(Year == 2001, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2001 <- raw_stomach_contents %>% 
  filter(Year == 2001, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2001)

#merge dataframes
poll_2001 <- full_join(WN_poll_2001, F_poll_2001, by = "Prey_Name")

################
#Year: 2003
WN_poll_2003 <- raw_stomach_contents %>% 
  filter(Year == 2001, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2003 <- raw_stomach_contents %>% 
  filter(Year == 2003, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2003)

#merge dataframes
poll_2003 <- full_join(WN_poll_2003, F_poll_2003, by = "Prey_Name")

################
#Year: 2005
WN_poll_2005 <- raw_stomach_contents %>% 
  filter(Year == 2005, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2005 <- raw_stomach_contents %>% 
  filter(Year == 2005, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2005)

#merge dataframes
poll_2005 <- full_join(WN_poll_2005, F_poll_2005, by = "Prey_Name")


###############
#Year: 2001
WN_poll_2007 <- raw_stomach_contents %>% 
  filter(Year == 2007, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2007 <- raw_stomach_contents %>% 
  filter(Year == 2007, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2007)

#merge dataframes
poll_2007 <- full_join(WN_poll_2007, F_poll_2007, by = "Prey_Name")

###############
#Year: 2009
WN_poll_2009 <- raw_stomach_contents %>% 
  filter(Year == 2009, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2009 <- raw_stomach_contents %>% 
  filter(Year == 2009, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2009)

#merge dataframes
poll_2009 <- full_join(WN_poll_2009, F_poll_2009, by = "Prey_Name")

###############
#Year: 2001
WN_poll_2001 <- raw_stomach_contents %>% 
  filter(Year == 2001, Pred_nodc == 8791030701) %>% #filter for year and Walleye Pollock
  group_by(Prey_Name) %>%  #group by prey 
  summarise(TotalWt = sum(Prey_twt), N = n()) %>% 
  mutate("%W" = (TotalWt/(sum(TotalWt))*100), "%N" = (N/sum(N))*100)

F_poll_2001 <- raw_stomach_contents %>% 
  filter(Year == 2001, Pred_nodc == 8791030701) %>% 
  group_by(Prey_Name) %>% 
  summarise(F = sum(unique(Pred_specn)*0+1)) %>% 
  mutate("%F" = (F/540)*100, year = 2001)

#merge dataframes
poll_2001 <- full_join(WN_poll_2001, F_poll_2001, by = "Prey_Name")

```

